---
title: "R Notebook"
output: R Workshop - Session 3
---
Conditional Statements

We use if, if-else-if or switch for conditional processes.

```{r}

if (value1 > value2) {
  cat(value1, " is greater than ", value2)
}

#If there is no output, it means value2 is greater than value1

```

```{r}

if (value1 > value2) {
  print("Still not greater.")
} else {
  print("The else block was executed here.")
}

```

```{r}

# Enter any weather type out of "Sunny", "Rainy", "Cold"
weather <- 

if (weather == "Sunny") {
  print("It is sunny out.")
} else if (weather == "Rainy") {
  print("It is raining")
} else {
  print("It is cold.")
}

```

```{r}

rollno <- 2

switch(3, "First", "Second", "Third", "Fourth")

value <- "34"

switch(value, "34"=print("The value is 34"), "46"=print("The value is 46"))

```

-------------------------

Q - WAP to check if temperature is above 25 degrees celcius. If it is above 25, print "It is mild.", else print "It is cold".

Q - WAP using a switch statement to print the day of the week.

Q - WAP which checks if the 5th value of a vector is greater than 10.

-------------------------

Other Stuff

'?function_name' to pull up the documentation of a function.

help.search('term') to search for any function with term in it.

help(package = 'package_name') to search for documentation of any package.

str(object_name) to get a summary of an object structure.

class(dataset_name) to know class of an object.

-----------------------

install.packages('package_name') to install a package.

library('package_name') to load package into the session.

package_name::function_name to use a particular function from a package.

data() shows all the available datasets in R.
data(dataset_name) to load a dataset into session.

------------------------

getwd() to get current working directory.

setwd('new_directory') to set working directory.

------------------------

More on Vectors

seq(start, end, skip) : Create a sequence

```{r}
print(seq(2, 10, 2))
print(seq(1, 100, 30))
```

rep(vector, times, each) : Repeat a vector

```{r}
rep(1:4, 3)
rep(c('a', 'b', 'c'), each=5)
rep(c(2, 4, 5), c(1, 4, 6))
```

sort(vector) : Sort a vector

```{r}
print(sort(c(100, 45, 23, 67)))
```

rev(vector) : Reverse a vector

```{r}
rev(c(9, 10, 24, 100, 98))
```

table(vector) : Count of each value

```{r}
print(table(c(23, 45, 23, 23, 89, 45)))
```

unique(vector) : return unique values from a vector

```{r}
print(unique(c(23, 45, 67, 23, 45)))
```

Loops

while loop

```{r}

x <- 0

while(x < 5) {
  print("Hello")
  x <- x + 1
}
```

for loop

```{r}

for (i in c(1, 3, 4)) {
  print(i)
}

```

Loop Control Statements 

break

```{r}

for (i in 1:10) {
  print(i)
  if (i == 5) {
    break
  }
}

for (i in 1:10) {
  if (i == 5) {
    break
  }
  print(i)
}

```

next 

```{r}

for (i in 1:20) {
  if ( i%%5 == 0 || i%%3 == 0 ) {
    next
  }
  print(i)
}

```

Strings 

paste(strings, sep=value)

```{r}

print(paste("Hello", "How are you?"))
print(paste("Hello", "How are you?", "What is your name?", sep="_"))

```

paste(vector, collapse)

```{r}
print(paste(c("Hello", "How", "are"), collapse = "+"))
```

strrep() : repeat a string N times

```{r}

strrep("Hello", 5)

```


Q. WAP to print all the numbers between 1 and 100 which are divisible by 5 or 7.
Q. WAP to print the following patterns:

*
**
***
****
*****

Q. WAP to print only the even numbers from a range.
Q. WAP to print the Fibonacci sequence.
Q. WAP to check which letters in a word are vowels.
Q. WAP to print factorial of a number.
Q. WAP to print the sum of the following series upto n: 
1 + 1/2 + 1/3 + 1/4 + ... + 1/n