---
title: "R Workshop Session 2"
output: html_notebook
---

Mathematical Operators

R is capable of performing basic mathematical operations, such as addition (+), multiplication (*), subtraction (-), division (/), quotient division (%/%), remainder division (%%), and exponentiation (^) on individual values or vectors.

```{r}

value1 <-  # Enter a value here
value2 <-  # Enter another value here
  
result.addition <- value1 + value2
print(result.addition)

result.multiplication <- value1 * value2
print(result.multiplication)

result.division <- value1 / value2
print(result.division)

result.subtraction <- value1 - value2
print(result.subtraction)

result.quotient <- value1 %/% value2
print(result.quotient)

result.remainder <- value1 %% value2
print(result.remainder)

result.exponent <- value1 ^ value2
print(result.exponent)

```

-------------------------

Performing operations on vectors

```{r}

vector1 <- #Add values here 
vector2 <- #Add values here

result.vector.addition <- vector1 + vector2 
print(result.vector.addition)

result.vector.multiplication <- vector1 * vector2
print(result.vector.multiplication)

result.vector.division <- vector1 / vector2
print(result.vector.division)

result.vector.subtraction <- vector1 - vector2
print(result.vector.subtraction)

result.vector.quotient <- vector1 %/% vector2
print(result.vector.quotient)

result.vector.remainder <- vector1 %% vector2
print(result.vector.remainder)

result.vector.exponent <- vector1 ^ vector2
print(result.vector.exponent)

```

-------------------------

Relational Operators

Relational Operators result in True or False.

There are 6 operators 

'>' : Greater than
'<' : Lesser than
'= =' : Equal to
'< =' : Less than or equal to
'> =' : Greater than or equal to
'! =' : Not equal to

```{r}

print(value1 > value2)
print(value1 < value2)
print(value1 == value2)
print(value1 >= value2)
print(value1 <= value2)
print(value1 != value2)

print(vector1 > vector2)
print(vector1 < vector2)
print(vector1 == vector2)
print(vector1 >= vector2)
print(vector1 <= vector2)
print(vector1 != vector2)

```

-------------------------

Logical Operators

There are 5 widely used logical operators

'&' : Element Wise Logical AND Operator
'&&' : Logical AND Operator (compares first element only)
'|' : Element Wise Logical OR Operator
'||' : Logical OR Operator (compares first element only)
'!' : Logical NOT Operator

```{r}

#print(value1 & value2)
#print(value1 | value2)
#print(!value1)

print(TRUE & FALSE)
print(TRUE & TRUE)
print(FALSE & FALSE)

print(TRUE | TRUE)
print(TRUE | FALSE)
print(FALSE | FALSE)

#print(vector1 & vector2)
#print(vector1 | vector2)
#print(vector1 || vector2)
#print(vector1 && vector2)
#print(!vector2)

```

-------------------------

Assignment Operators

Most widely used assignment operator is '< -', but there are few more operators:

'< -', '<< -", "=" called Left Hand Assignment
'- >', '- >>' called Right Hand Assignment

-------------------------

Other Operators

':' Creates a vector or series of numbers starting from the value on its left to the value on its right.

'%in%' checks if a value is in a vector/list/matrix

'%*%' Multiplies two matrices
```{r}

print(1:4)

1 %in% 1:4

```
```{r}

a <- 4

A <- 5

Economics
economics

print(a)
print(A)

```



```{r}

series.vector <- # Create a series

check_value <-  # Enter any value
  
print(check_value %in% series.vector)

test.matrix <- # Create a matrix

# Multiply this matrix by its transpose. To transpose a matrix M, we use the function t(M).
transpose_product <- 
print(transpose_product)
  
```

WAP which multiplies a 4x5 matrix with its transpose.

```{r}

```

-------------------------

Printing

print() function can print only one argument at once. To print multiple arguments, we use the cat() function.

```{r}

cat(value1, value2)

# This will return an error
print(value1, value2)

```

-------------------------

Data type of a variable

We use the class() function to check the data type of a variable.

-------------------------

Selecting Elements from a vector.

```{r}

vector1 <- c(3, 6, 8, 10, 15, 20, 21)

# Printing first value in vector1.
print(vector1[1])

# Print all elements but the first element from vector1.
print(vector1[-1])

# Print elements 2 to 4 from vector1
print(vector1[2:4])

# Print all elements except 2 to 4 from vector1.
print(vector1[-(2:4)])

# Print Elements 1 and 3
print(vector1[c(1,3)])

# Print all elements which are equal to 10.
print(vector1[vector1 == 10])

# Print all elements less than 10
print(vector1[vector1 < 10])

# Print elements in a particular set
print(vector1[vector1 %in% c(3, 4, 10)])

```

-------------------------

Selecting Elements from Matrix

```{r}

M <- #create a 4x4 matrix
  
# Selecting a row

#First row
M[1, ]

#First Column
M[ ,1]

#Selecting an element
M[1,1]

# Solve a matrix equation

N <- #create another 4x3 matrix
# Solve MX = N for X
solve(M, N)

```

-------------------------

Conditional Statements

We use if, if-else-if or switch for conditional processes.

```{r}

if (value1 > value2) {
  cat(value1, " is greater than ", value2)
}

#If there is no output, it means value2 is greater than value1

```

```{r}

if (value1 > value2) {
  print("Still not greater.")
} else {
  print("The else block was executed here.")
}

```

```{r}

# Enter any weather type out of "Sunny", "Rainy", "Cold"
weather <- 

if (weather == "Sunny") {
  print("It is sunny out.")
} else if (weather == "Rainy") {
  print("It is raining")
} else {
  print("It is cold.")
}

```

```{r}

rollno <- 2

switch(3, "First", "Second", "Third", "Fourth")

value <- "34"

switch(value, "34"=print("The value is 34"), "46"=print("The value is 46"))

```

-------------------------

Q - WAP to check if temperature is above 25 degrees celcius. If it is above 25, print "It is mild.", else print "It is cold".

Q - WAP using a switch statement to print the day of the week.

Q - WAP which checks if the 5th value of a vector is greater than 10.

-------------------------


